Score-based tests of measurement invariance: Use in practice

### 2013-09-05 frontiers/frontiers.Rnw
go through the intro,tutorial and sim method sections, change some typos and write some small questions.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@248 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Following model estimation, the tests primarily work on the {\em
  scores} of the fitted model; these are defined as
%% TODO indicate that this is evaluated at theta hat
\begin{equation}
\label{eq:score}
s(\hat{\theta};x_{i}) = \left ( \frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{k})} \right )^{T},
\end{equation} 
where $\ell(\theta; x_i)$ is the likelihood associated with individual
$i$, $\theta$ is a $k$-dimensional parameter vector, and
$\hat{\theta}$ is the maximum likelihood estimate.  The cross-product
of these scores forms the ``meat'' for the calculation of robust
(Huber-White) standard errors. %TODO cite

@@

We focus on measurement invariance of the factor loadings associated with t
he GQ-6 scale, using a one-factor model.While the ``age group'' variable against 
which we are testing measurement invariance is best considered ordinal, 
for demonstration we consider its treatment as categorical, continuous, and ordinal.  
Each of these treatments is described below in a separate section.

@@

\fixme{Need to update the citation below so that the last name is ``van de Schoot''.}
Measurement invariance is most often tested using multiple groups models \cite<see, e.g., >{vande12}.  
This amounts to assuming that our auxiliary variable is categorical (i.e., unordered), 
which is not true for the age groups in the data.  However, we demonstrate the procedure for completeness.

@@

\subsection{Continuous Treatment}
If we are interested in measurement invariance violations that are 
monotonic with the age groups, it is perhaps simplest to treat the 
age groups as continuous.  In doing so, we can use the statistics 
from Equations~\eqref{eq:dmax},~\eqref{eq:cvm}, and~\eqref{eq:maxlm}.  
That is, we can fit a model whose parameters 
are restricted to be equal across across all individuals and then examine 
how individuals' scores $s(\hat{theta};x_{i})$ fluctuate with their
age (where age ties are broken arbitrarily).
This is demonstrated below, with similar code being useful when one is 
testing measurement invariance w.r.t.\ other continuous variables.

@@

Power and Type I error were examined across three sample sizes
($n=120,480,960$) , three numbers of categories ($m=4,8,12$)and
7 magnitudes of invariance violations. 
The meaurement invariance violations began at level $1+m/2$ of $V$
and were consistant thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,1.5$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested.Four statistics were examined above. In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

@@

\section{Simulation 2}
In Simulation 2, we examine the extent to which the results of
Simulation 1 are robust to model misspecification. Specifically, we
generate data from the factor analysis model used in the previous
section, but with adding an extra loading from the second factor to the first
indicator (of the first factor).  The estimated model lacks this extra
loading, however. The goal of this simulation is to examine the proposed statitics'
power and spicificity under model misspecification.

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra,unmodeled loading. 
Sample size and magnitude of measurementinvariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

>>>>

Following model estimation, the tests primarily work on the {\em
  scores} of the fitted model; these are defined as
%% TODO indicate that this is evaluated at theta hat
\begin{equation}
\label{eq:score}
s(\hat{\theta};x_{i}) = \left ( \frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial
      \ell(\theta;x_{i})}{\partial(\theta_{k})} \right )^{T},
\end{equation} 
where $\ell(\theta; x_i)$ is the likelihood associated with individual
$i$, $\theta$ is a $k$-dimensional parameter vector, and
$\hat{\theta}$ is the maximum likelihood estimate.  The cross-product
of these scores forms the ``meat'' for the calculation of robust
(Huber-White) standard errors. %TODO cite
\seeme{I don't understand this sentence. cross-product of each individual for each parameter? Why this 
is the standard error?}

@@

We focus on measurement invariance of the factor loadings associated with 
the GQ-6 scale, using a one-factor model. While the ``age group'' variable against 
which we are testing measurement invariance is best considered ordinal, 
for demonstration we consider its treatment as categorical, continuous, and ordinal.  
Each of these treatments is described below in a separate section.

@@

\fixme{Need to update the citation below so that the last name is ``van de Schoot''.}
\seeme{Is it a different paper from vande12?}
Measurement invariance is most often tested using multiple groups models \cite<see, e.g., >{vande12}.  
This amounts to assuming that our auxiliary variable is categorical (i.e., unordered), 
which is not true for the age groups in the data.  However, we demonstrate the procedure for completeness.

@@

\subsection{Continuous Treatment}
If we are interested in measurement invariance violations that are 
monotonic with the age groups, it is perhaps simplest to treat the 
age groups as continuous.  In doing so, we can use the statistics 
from Equations~\eqref{eq:dmax},~\eqref{eq:cvm}, and~\eqref{eq:maxlm}.  
That is, we can fit a model whose parameters 
are restricted to be equal across all individuals and then examine 
how individuals' scores $s(\hat{theta};x_{i})$ fluctuate with their
age (where age ties are broken arbitrarily).
This is demonstrated below, with similar code being useful when one is 
testing measurement invariance w.r.t.\ other continuous variables.

@@

Power and Type I error were examined across three sample sizes
($n=120,480,960$) , three numbers of categories ($m=4,8,12$)and
17 magnitudes of invariance violations. 
The measurement invariance violations began at level $1+m/2$ of $V$
and were consistent thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested. Four statistics were examined above. In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

@@

\section{Simulation 2}
In Simulation 2, we examine the extent to which the results of
Simulation 1 are robust to model misspecification. Specifically, we
generate data from the factor analysis model used in the previous
section, but with adding an extra loading from the second factor to the first
indicator (of the first factor).  The estimated model lacks this extra
loading, however. The goal of this simulation is to examine the proposed statistics'
power and specificity under model misspecification.

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra, unmodeled loading. 
Sample size and magnitude of measurementinvariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.
